---
  
  ### **R Markdown File - Null Model Randomization within Localities**
  ```r
---
  title: "Null Model Randomization by Locality"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---
  
  ```{r setup, include=FALSE}
# Load required libraries
library(vegan)
library(dplyr)
```

## **Introduction**
In this tutorial, we apply the **quasiswap null model** separately within each locality to maintain row and column sums while randomizing data.

## **Step 1: Generate Example Data**
```{r}
set.seed(123)
localities <- rep(c("A", "B", "C"), each = 10)  # Locality labels
data <- data.frame(Locality = localities, matrix(sample(1:100, 50, replace = TRUE), nrow = 30, ncol = 5))

# Print first few rows
head(data)
```

## **Step 2: Apply Quasiswap Randomization Per Locality**
```{r}
randomize_by_locality <- function(df) {
  df %>%
    group_by(Locality) %>%
    group_modify(~ data.frame(Locality = .x$Locality, nullmodel(as.matrix(.x[, -1]), method = "quasiswap"))) %>%
    ungroup()
}

null_data <- randomize_by_locality(data)
head(null_data)
```

## **Step 3: Compare Observed and Null Model Data**
```{r}
observed_mean <- mean(as.matrix(data[, -1]))
null_mean <- mean(as.matrix(null_data[, -1]))

cat("Observed Mean:", observed_mean, "\n")
cat("Null Model Mean:", null_mean, "\n")
```

## **Step 4: Visualize the Data**
```{r}
par(mfrow = c(1,2))
image(as.matrix(data[data$Locality == "A", -1]), main = "Observed Data (Locality A)", col = heat.colors(20))
image(as.matrix(null_data[null_data$Locality == "A", -1]), main = "Null Model Data (Locality A)", col = heat.colors(20))
```


